### 2012-03-31 doc/theses/mikula/appendix_a.tex
Master thesis: In progress.

\chapter{Attachments}
\begin{figure}
\caption{All possible values representing an operator in an instance of \texttt{OperatorNode}}
\label{FIG_operators}
Logical: \texttt{AND}, \texttt{OR}.

@@

Unary: \texttt{UNARY\_PLUS}, \texttt{UNARY\_MINUS}.
\end{figure}

\begin{figure}
\caption{XSD built-in atomic types.}
\label{FIG_xsd_built_in_atomic_types}
\includegraphics[scale=0.65]{xsd-types.eps}
\end{figure}

>>>>

\chapter{Attachments}
\begin{figure}[float=htpb]
Logical: \texttt{AND}, \texttt{OR}.

@@

Unary: \texttt{UNARY\_PLUS}, \texttt{UNARY\_MINUS}.
\caption{All possible values representing an operator in an instance of \texttt{OperatorNode}}
\label{FIG_operators}
\end{figure}

\begin{lstlisting}[caption=Resulting XSD of Test Scenario B, frame=single, label=listing_xsd]
<?xml version="1.0" encoding="UTF-8"?>
<xs:schema xmlns:xs="http://www.w3.org/2001/XMLSchema">
<!-- Inferred on Sat Mar 31 21:28:32 CEST 2012 by Basic IG generator, TwoStep(Iname (with attributes), Automaton Merging State(GreedyMDL(Combined(k,h-context, s,k-strings, Null, Null),Naive Alphabet), State Removal Ordered(Weighted)), Chained(Empty Children, Nested Concatenation, Null)), Basic XSD exporter -->

<!-- global types -->
<xs:complexType name="Temph" mixed="true">
  <xs:sequence>
    <xs:sequence minOccurs="0" maxOccurs="unbounded">
      <xs:choice>
        <xs:element name="keyword" type="Tkeyword"/>
        <xs:element name="bold" type="Tbold"/>
      </xs:choice>
    </xs:sequence>
  </xs:sequence>
</xs:complexType>

<xs:complexType name="Tbold" mixed="true">
  <xs:sequence>
    <xs:choice>
      <xs:choice>
        <xs:sequence>
          <xs:element name="emph" type="Temph"/>
        </xs:sequence>
      </xs:choice>
      <xs:sequence>
        <xs:element name="keyword" type="Tkeyword"/>
      </xs:sequence>
    </xs:choice>
  </xs:sequence>
</xs:complexType>

<xs:complexType name="Tkeyword" mixed="true">
  <xs:sequence>
    <xs:choice>
      <xs:sequence minOccurs="0">
        <xs:element name="bold" type="Tbold"/>
      </xs:sequence>
      <xs:sequence>
        <xs:element name="emph" type="Temph"/>
      </xs:sequence>
    </xs:choice>
  </xs:sequence>
</xs:complexType>

<xs:complexType name="Ttext" mixed="true">
  <xs:choice minOccurs="0" maxOccurs="unbounded">
    <xs:element name="bold" type="Tbold"/>
    <xs:element name="keyword" type="Tkeyword"/>
    <xs:element name="emph" type="Temph"/>
  </xs:choice>
</xs:complexType>

<xs:complexType name="Tparlist">
  <xs:sequence>
    <xs:element name="listitem" type="Tlistitem" minOccurs="0" maxOccurs="unbounded"/>
  </xs:sequence>
</xs:complexType>

<xs:complexType name="Tlistitem">
  <xs:choice>
    <xs:element name="text" type="Ttext"/>
    <xs:element name="parlist" type="Tparlist"/>
  </xs:choice>
</xs:complexType>

<xs:complexType name="Tdescription">
  <xs:choice>
    <xs:element name="parlist" type="Tparlist"/>
    <xs:element name="text" type="Ttext"/>
  </xs:choice>
</xs:complexType>

<xs:complexType name="Tincategory">
  <xs:attribute name="category" type="xs:string" use="required"/>
</xs:complexType>

<xs:complexType name="Tmail">
  <xs:sequence>
    <xs:element name="from" type="xs:string"/>
    <xs:element name="to" type="xs:string"/>
    <xs:element name="date" type="xs:string"/>
    <xs:element name="text" type="Ttext"/>
  </xs:sequence>
</xs:complexType>

<xs:complexType name="Tmailbox">
  <xs:sequence minOccurs="0">
    <xs:element name="mail" type="Tmail"/>
    <xs:element name="mail" type="Tmail" minOccurs="0"/>
  </xs:sequence>
</xs:complexType>

<xs:complexType name="Titem">
  <xs:sequence>
    <xs:element name="location" type="xs:string"/>
    <xs:element name="quantity" type="xs:string"/>
    <xs:element name="name" type="xs:string"/>
    <xs:element name="payment" type="xs:string"/>
    <xs:element name="description" type="Tdescription"/>
    <xs:element name="shipping" type="xs:string"/>
    <xs:element name="incategory" type="Tincategory" minOccurs="0" maxOccurs="unbounded"/>
    <xs:element name="mailbox" type="Tmailbox"/>
  </xs:sequence>
  <xs:attribute name="id" type="xs:string" use="required"/>
</xs:complexType>

<xs:complexType name="Tafrica">
  <xs:sequence>
    <xs:element name="item" type="Titem"/>
  </xs:sequence>
</xs:complexType>

<xs:complexType name="Tasia">
  <xs:sequence>
    <xs:element name="item" type="Titem"/>
    <xs:element name="item" type="Titem"/>
  </xs:sequence>
</xs:complexType>

<xs:complexType name="Taustralia">
  <xs:sequence>
    <xs:element name="item" type="Titem"/>
    <xs:element name="item" type="Titem"/>
  </xs:sequence>
</xs:complexType>

<xs:complexType name="Teurope">
  <xs:sequence>
    <xs:element name="item" type="Titem" minOccurs="0" maxOccurs="unbounded"/>
  </xs:sequence>
</xs:complexType>

<xs:complexType name="Tnamerica">
  <xs:sequence>
    <xs:element name="item" type="Titem" minOccurs="0" maxOccurs="unbounded"/>
  </xs:sequence>
</xs:complexType>

<xs:complexType name="Tsamerica">
  <xs:sequence>
    <xs:element name="item" type="Titem"/>
  </xs:sequence>
</xs:complexType>

<xs:complexType name="Tregions">
  <xs:sequence>
    <xs:element name="africa" type="Tafrica"/>
    <xs:element name="asia" type="Tasia"/>
    <xs:element name="australia" type="Taustralia"/>
    <xs:element name="europe" type="Teurope"/>
    <xs:element name="namerica" type="Tnamerica"/>
    <xs:element name="samerica" type="Tsamerica"/>
  </xs:sequence>
</xs:complexType>

<xs:complexType name="Tcategory">
  <xs:sequence>
    <xs:element name="name" type="xs:string"/>
    <xs:element name="description" type="Tdescription"/>
  </xs:sequence>
  <xs:attribute name="id" type="xs:string" use="required"/>
</xs:complexType>

<xs:complexType name="Tcategories">
  <xs:sequence>
    <xs:element name="category" type="Tcategory"/>
  </xs:sequence>
</xs:complexType>

<xs:complexType name="Tedge">
  <xs:attribute name="from" type="xs:string" use="required"/>
  <xs:attribute name="to" type="xs:string" use="required"/>
</xs:complexType>

<xs:complexType name="Tcatgraph">
  <xs:sequence>
    <xs:element name="edge" type="Tedge"/>
  </xs:sequence>
</xs:complexType>

<xs:complexType name="Taddress">
  <xs:sequence>
    <xs:element name="street" type="xs:string"/>
    <xs:element name="city" type="xs:string"/>
    <xs:element name="country" type="xs:string"/>
    <xs:choice>
      <xs:element name="zipcode" type="xs:string"/>
      <xs:sequence>
        <xs:element name="province" type="xs:string"/>
        <xs:element name="zipcode" type="xs:string"/>
      </xs:sequence>
    </xs:choice>
  </xs:sequence>
</xs:complexType>

<xs:complexType name="Tinterest">
  <xs:attribute name="category" type="xs:string" use="required"/>
</xs:complexType>

<xs:complexType name="Tprofile">
  <xs:sequence>
    <xs:choice minOccurs="0" maxOccurs="unbounded">
      <xs:element name="education" type="xs:string"/>
      <xs:element name="interest" type="Tinterest"/>
    </xs:choice>
    <xs:choice>
      <xs:element name="business" type="xs:string"/>
      <xs:sequence>
        <xs:element name="gender" type="xs:string"/>
        <xs:element name="business" type="xs:string"/>
      </xs:sequence>
    </xs:choice>
    <xs:element name="age" type="xs:string" minOccurs="0"/>
  </xs:sequence>
  <xs:attribute name="income" type="xs:integer" use="required"/>
</xs:complexType>

<xs:complexType name="Twatch">
  <xs:attribute name="open_auction" type="xs:string" use="required"/>
</xs:complexType>

<xs:complexType name="Twatches">
  <xs:sequence>
    <xs:element name="watch" type="Twatch" minOccurs="0" maxOccurs="unbounded"/>
  </xs:sequence>
</xs:complexType>

<xs:complexType name="Tperson">
  <xs:sequence>
    <xs:element name="name" type="xs:string"/>
    <xs:element name="emailaddress" type="xs:string"/>
    <xs:choice minOccurs="0" maxOccurs="unbounded">
      <xs:element name="address" type="Taddress"/>
      <xs:element name="phone" type="xs:string"/>
      <xs:element name="homepage" type="xs:string"/>
      <xs:element name="creditcard" type="xs:string"/>
    </xs:choice>
    <xs:choice>
      <xs:sequence minOccurs="0">
        <xs:element name="profile" type="Tprofile"/>
        <xs:element name="watches" type="Twatches" minOccurs="0"/>
      </xs:sequence>
      <xs:element name="watches" type="Twatches"/>
    </xs:choice>
  </xs:sequence>
  <xs:attribute name="id" type="xs:string" use="required"/>
</xs:complexType>

<xs:complexType name="Tpeople">
  <xs:sequence>
    <xs:element name="person" type="Tperson" minOccurs="0" maxOccurs="unbounded"/>
  </xs:sequence>
</xs:complexType>

<xs:complexType name="Tpersonref">
  <xs:attribute name="person" type="xs:string" use="required"/>
</xs:complexType>

<xs:complexType name="Tbidder">
  <xs:sequence>
    <xs:element name="date" type="xs:string"/>
    <xs:element name="time" type="xs:string"/>
    <xs:element name="personref" type="Tpersonref"/>
    <xs:element name="increase" type="xs:integer"/>
  </xs:sequence>
</xs:complexType>

<xs:complexType name="Titemref">
  <xs:attribute name="item" type="xs:string" use="required"/>
</xs:complexType>

<xs:complexType name="Tseller">
  <xs:attribute name="person" type="xs:string" use="required"/>
</xs:complexType>

<xs:complexType name="Tauthor">
  <xs:attribute name="person" type="xs:string" use="required"/>
</xs:complexType>

<xs:complexType name="Tannotation">
  <xs:sequence>
    <xs:element name="author" type="Tauthor"/>
    <xs:element name="description" type="Tdescription"/>
    <xs:element name="happiness" type="xs:string"/>
  </xs:sequence>
</xs:complexType>

<xs:complexType name="Tinterval">
  <xs:sequence>
    <xs:element name="start" type="xs:string"/>
    <xs:element name="end" type="xs:string"/>
  </xs:sequence>
</xs:complexType>

<xs:complexType name="Topen_auction">
  <xs:sequence>
    <xs:element name="initial" type="xs:integer"/>
    <xs:choice minOccurs="0" maxOccurs="unbounded">
      <xs:element name="reserve" type="xs:decimal"/>
      <xs:element name="bidder" type="Tbidder"/>
    </xs:choice>
    <xs:element name="current" type="xs:string"/>
    <xs:element name="privacy" type="xs:string" minOccurs="0" maxOccurs="unbounded"/>
    <xs:element name="itemref" type="Titemref"/>
    <xs:element name="seller" type="Tseller"/>
    <xs:element name="annotation" type="Tannotation"/>
    <xs:element name="quantity" type="xs:string"/>
    <xs:element name="type" type="xs:string"/>
    <xs:element name="interval" type="Tinterval"/>
  </xs:sequence>
  <xs:attribute name="id" type="xs:string" use="required"/>
</xs:complexType>

<xs:complexType name="Topen_auctions">
  <xs:sequence>
    <xs:element name="open_auction" type="Topen_auction" minOccurs="0" maxOccurs="unbounded"/>
  </xs:sequence>
</xs:complexType>

<xs:complexType name="Tbuyer">
  <xs:attribute name="person" type="xs:string" use="required"/>
</xs:complexType>

<xs:complexType name="Tclosed_auction">
  <xs:sequence>
    <xs:element name="seller" type="Tseller"/>
    <xs:element name="buyer" type="Tbuyer"/>
    <xs:element name="itemref" type="Titemref"/>
    <xs:element name="price" type="xs:integer"/>
    <xs:element name="date" type="xs:string"/>
    <xs:element name="quantity" type="xs:string"/>
    <xs:element name="type" type="xs:string"/>
    <xs:element name="annotation" type="Tannotation"/>
  </xs:sequence>
</xs:complexType>

<xs:complexType name="Tclosed_auctions">
  <xs:sequence>
    <xs:element name="closed_auction" type="Tclosed_auction" minOccurs="0" maxOccurs="unbounded"/>
  </xs:sequence>
</xs:complexType>

<xs:complexType name="Tsite">
  <xs:sequence>
    <xs:element name="regions" type="Tregions"/>
    <xs:element name="categories" type="Tcategories"/>
    <xs:element name="catgraph" type="Tcatgraph"/>
    <xs:element name="people" type="Tpeople"/>
    <xs:element name="open_auctions" type="Topen_auctions"/>
    <xs:element name="closed_auctions" type="Tclosed_auctions"/>
  </xs:sequence>
</xs:complexType>

<!-- top level element -->
<xs:element name="site" type="Tsite">
  <xs:key name="key1">
    <xs:selector xpath="people/person"/>
    <xs:field xpath="@id"/>
  </xs:key>
  <xs:key name="key2">
    <xs:selector xpath=".//item"/>
    <xs:field xpath="@id"/>
  </xs:key>
  <xs:key name="key3">
    <xs:selector xpath="regions/europe/item"/>
    <xs:field xpath="@id"/>
  </xs:key>
  <xs:keyref name="keyRef1_key1" refer="key1">
    <xs:selector xpath="closed_auctions/closed_auction"/>
    <xs:field xpath="buyer/@person"/>
  </xs:keyref>
  <xs:keyref name="keyRef2_key2" refer="key2">
    <xs:selector xpath="closed_auctions/closed_auction"/>
    <xs:field xpath="itemref/@item"/>
  </xs:keyref>
  <xs:keyref name="keyRef3_key3" refer="key3">
    <xs:selector xpath="closed_auctions/closed_auction"/>
    <xs:field xpath="itemref/@item"/>
  </xs:keyref>
  <xs:keyref name="keyRef4_key3" refer="key3">
    <xs:selector xpath="open_auctions/open_auction"/>
    <xs:field xpath="itemref/@item"/>
  </xs:keyref>
</xs:element>
</xs:schema>
\end{lstlisting}

### 2012-03-31 doc/theses/mikula/experiments.tex
Master thesis: In progress.

\section{Test Data}
To get meaningful results of the experiments, test data should be composed of XML documents, which are instances of a certain, possibly not known, XML schema, and a set of XQuery queries which query the XML documents. The amount of the XML data does not have to be large. On the other hand, the set of queries should be large (at least hundreds of queries) and the queries should be real, not artificially made.

@@

\section{Results}

\subsection{Type Inference}
Six type statements shown in Table \ref{TAB_inferred_types} were inferred.

@@

\subsection{Key Discovery}
Four key statements and their normalized weights shown in Table \ref{TAB_inferred_keys} were inferred.

@@

\todo[inline]{Pripojit vysledne XSD? Je dost velke. Alebo aspon relevantnu cast? A: Neni treba, kdyz tak do appendixu.}

>>>>

We made two test scenarios; one dealing with input data that were not made for purposes of this experiments, and the other extending the first dataset by data created for better test coverage.

\section{Test Scenario A}

\subsection{Test Data}
To get meaningful results of the experiments, test data should be composed of XML documents, which are instances of a certain, possibly not known, XML schema, and a set of XQuery queries which query the XML documents. The amount of the XML data does not have to be large. On the other hand, the set of queries should be large (at least hundreds of queries) and the queries should be real, not artificially made.

@@

\subsection{Results}

\subsubsection{Type Inference}
Six type statements shown in Table \ref{TAB_inferred_types} were inferred.

@@

\subsubsection{Key Discovery}
Four key statements and their normalized weights shown in Table \ref{TAB_inferred_keys} were inferred.

@@

\section{Test Scenario B}
As mentioned before, the previous test scenario involves only data with JP3 occurrences, and therefore, it only tests the extension of the original method, while the original method itself remains untested. To correct it, we add several new test queries created by ourselves.

\subsection{Test Data}

\begin{lstlisting}[caption=Test query B1 containing a for join pattern occurrence., frame=single, label=listing_test_query_b1]
for $item in /site/regions/europe/item
let $closed_auctions := /site/closed_auctions/closed_auction[itemref/@item = $item/@id]
return <item><id>{$item/@id}</id><max-price>{max($closed_auctions/price)}</max-price></item>
\end{lstlisting}

\begin{lstlisting}[caption=Test query B2 containing a let join pattern occurrence., frame=single, label=listing_test_query_b2]
for $open_auction in /site/open_auctions/open_auction
let $item := /site/regions/europe/item[@id = $open_auction/itemref/@item]
return $item/personref/@person
\end{lstlisting}

\begin{lstlisting}[caption=Test query B3 containing let join pattern and JP3 occurrences., frame=single, label=listing_test_query_b3, numbers=left, numberstyle=\tiny]
for $person in /site/people/person
where $person/profile/@income < 100.00 and $person/profile/gender = "m"
return
    <list>{
        for $auction in /site/closed_auctions/closed_auction
        let $item := /site//item[@id = $auction/itemref/@item]
        let $price := $auction/price
        where $auction/buyer/@person = $person/@id
        return
            <record><person>{$person/@id}</person><item>{$item/@id}</item><price>{$price}</price></record>
    }</list>
\end{lstlisting}

In this scenario, we use test data from the Test Scenario A extended by queries in Listings \ref{listing_test_query_b1}, \ref{listing_test_query_b2}, and \ref{listing_test_query_b3}.

\todo[inline]{Mozno popisat aj vyznam tychto dotazov?}

The first two ones are simple queries containing one join pattern occurrence each. The third one contains occurrences of let join pattern (binding clauses at lines 5 and 6) and JP3 (binding clauses at lines 1 and 5, and where clause at line 8), two expressions exploitable by the inference of types (both at line 2) and one negative uniqueness statement of the comparison-with-a-constant form (comparison with \texttt{"m"} at line 2). Thus, the third query contains instances of all types of the constructs utilized by our method.

\subsubsection{Results}

\begin{table}
\begin{tabular}{|l|}
\hline
\texttt{/site/open\_auctions/open\_auction/bidder/increase/text()} $\rightarrow$ \texttt{integer} \\ \hline
\texttt{/site/closed\_auctions/closed\_auction/price/text()} $\rightarrow$ \texttt{integer} \\ \hline
\texttt{/site/open\_auctions/open\_auction/initial/text()} $\rightarrow$ \texttt{integer} \\ \hline
\texttt{/site/open\_auctions/open\_auction/initial/text()} $\rightarrow$ \texttt{integer} \\ \hline
\texttt{/site/people/person/profile/@income} $\rightarrow$ \texttt{integer} \\ \hline
\texttt{/site/open\_auctions/open\_auction/reserve} $\rightarrow$ \texttt{decimal} \\ \hline
\texttt{/site/people/person/profile/@income} $\rightarrow$ \texttt{decimal} \\ \hline
\texttt{/site/people/person/profile/gender} $\rightarrow$ \texttt{string} \\ \hline
\end{tabular}
\caption{Inferred type statements B.}
\label{TAB_inferred_types_b}
\end{table}

\begin{table}
\begin{tabular}{|l|c|}
\hline
\textbf{Key} & \textbf{Weight} \\ \hline \hline
\texttt{(/site//item, {@id})} & 0.333 \\ \hline
\texttt{(/site/people/person, {@id})} & 0.6 \\ \hline
\texttt{(/site/closed\_auctions/closed\_auction, {buyer/@person})} & -0.6 \\ \hline
\texttt{(/site/closed\_auctions/closed\_auction, {itemref/@item})} & -0.917 \\ \hline
\texttt{(/site/regions/europe/item, {@id})} & 0.6 \\ \hline
\end{tabular}
\caption{Inferred key statements B.}
\label{TAB_inferred_keys_b}
\end{table}

\begin{table}
\begin{tabular}{|l|c|}
\hline
\textbf{Node} & \textbf{Weight} \\ \hline \hline
\texttt{(/site/people/person/profile/interest/@category)} & 1 \\ \hline
\texttt{(/site/closed\_auctions/closed\_auction/itemref/@item)} & 1 \\ \hline
\texttt{(/site/people/person/profile/gender)} & 0.9 \\ \hline
\end{tabular}
\caption{Inferred negative uniqueness statements B.}
\label{TAB_inferred_negative_uniqueness_statements_b}
\end{table}

Results of this test scenario are shown in Tables \ref{TAB_inferred_types_b}, \ref{TAB_inferred_keys_b}, and \ref{TAB_inferred_negative_uniqueness_statements_b}. All seams to be as we expected.

Note that all inferred type statements are correct, however, the weight of the \texttt{(/site//item, {@id})} key is only 0.333. The reason is that there is only one join pattern occurrence resulting to this key, and therefore in summary, its normalized weight is low and it is correct.

We demonstrated that also the original method works and we can get better results with little larger input dataset. Though, this set is still very small and we do not show how the methods works with large real-world data, possibly containing FLWOR constructs not satisfying our assumption that each join is done by a key/foreign key pair.

We attach the resulting XML Schema in Appendix \ref{listing_xsd}. It was made using threshold 0.3. It means that all key statements with normalized weight equal or higher than 0.3 are considered correct and are included in the schema.
